### 2021-06-16 paper.Rmd
Old files.

---
title: |
  |
  | 
  | \vspace{1cm}Writing a reproducible paper
  | in Pagedown*^[*Corresponding address: mail@paulcbauer.eu. Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]\vspace{0.5cm}
  |
author: |
  | Paul C. Bauer & Camille Landesvatter
  | Mannheim Centre for European Social Research
date: |
  |       
  | Based on: https://osf.io/q395s
  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
  |
  | Feedback? Errors?
  | mail@paulcbauer.eu
  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded [here](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
bibliography: references.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin=1in
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use chache = TRUE if you want to speed up compilation

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

# Why reproducible research (in R)?
Some arguments...

* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].
* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years. 
* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.

# Prerequesites
I assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown. If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416). Then...

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex].
```{r eval=FALSE, include=TRUE}
install.packages(c('tinytex', 'rmarkdown'))
tinytex::install_tinytex()
```
* ...install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "plotly", "knitr",
                   "bookdown"))
```
* ...download the 5 input files I created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). Ignore the other files.

* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).

# Basics: Input and output files
All the files you need to produce the present PDF file are the input files...

* ...a `paper.rmd` file (the underlying R Markdown file).
* ...a `references.bib` file (the bibliography).
    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.rmd` file.
* ...a `data.csv` file (some raw data).
* ... a `american-sociological-association.csl` file that defines the style of your bibliography.^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]

[Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder. Close R/Rstudio and directly open `paper.rmd` with RStudio. Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]

Once you run/compile the `paper.rmd` file in Rstudio it creates mainly two output files:

* `paper.tex`
* `paper.pdf` (the one you are reading right now)

In addition, there may be files that you generate and store locally in the folder during the compilation process. This is the case for some of the Plotly graphs below.

Ideally, we can simply provide others with a `zip` folder that contains both our input files and our output files. Then it's possible to reproduce the process from managing/analyzing some raw data to producing the final scientific article.

Below we always display the R code in the chunks that produce the output. In your paper you will normally only present outputs (e.g., tables, figures etc.) by choosing the chunk option "Show output only" in R Studio. The chunk commands itself are not displayed but they do matter for referencing etc. So simply orient yourself at the underlying `paper.rmd` file.

# Referencing within your document

To see how referencing works simply see the different examples for figures, tables and sections below. For instance in Section \@ref(sec:tables) you can find different ways of referencing tables. The code of the underlying `paper.rmd` will show you how I referenced Section \@ref(sec:tables) right here namely with '`Section \@ref(sec:tables)`'.

# Software versioning
Software changes and gets updated, especially with an active developer community like that of R. Luckily you can always access [old versions of R](https://cran.r-project.org/bin/windows/base/old/) and old version of R packages in [the archive](https://cran.r-project.org/src/contrib/Archive/). In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., `dplyr_0.2.tar.gz`. Then insert the path in the following function: `install.packages("https://....../dplyr_0.2.tar.gz", repos=NULL, type="source")`. Ideally, however, results will be simply reproducible in the most current R and package versions.

I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \@ref(sec:rsessioninfo). This will make sure you always provide the package versions that you used in the last compilation of your paper. For more advanced tools see [packrat](https://rstudio.github.io/packrat/).

```{r fig-versioning, echo=TRUE, eval=FALSE}
cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
  # or use message() instead of cat()
```

# Data
## Import
```{r, echo=TRUE}
data <- read.csv("data.csv")
head(data)
```

## Putting your entire data into the .rmd file
Applying the function `dput()` to an object gives you the code needed to reproduce that object. So you could paste that code into your `.rmd` file if you don't want to have extra data files. This makes sense were data files are small.

```{r}
dput(data)
```

You can then insert the dput output in your `.rmd` as below.

```{r eval=FALSE, include=TRUE}
data <- structure(list(X = 1:50, speed = c(4L, 4L, 7L, 7L, 8L, 9L, 10L, 
10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, 
14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, 
18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, 
24L, 24L, 24L, 25L), dist = c(2L, 10L, 4L, 22L, 16L, 10L, 18L, 
26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, 
36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, 
76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, 
92L, 93L, 120L, 85L)), 
class = "data.frame", row.names = c(NA, 
-50L))
```

# Tables {#sec:tables}
Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.

## stargazer(): Summary and regression tables

Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab1", 
          table.placement = "H", 
          header=FALSE)
```

Table \@ref(tab2) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
model1 <- lm(speed ~ dist, data = cars)
model2 <- lm(speed ~ dist, data = cars)
model3 <- lm(dist ~ speed, data = cars)
stargazer(model1, model2, model3,
          title = "Regression table with stargazer",
          label="tab2", 
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

## kable() and kable_styling()

Another great function is `kable()` (`knitr` package) in combination with `kableExtra`. Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.] Again you can modify so many things in both the `kable()` and the `kable_styling()` function. See [this overview](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) of all the kable stylings that are possible provided by the package author himself.

```{r tab3, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(knitr)
library(kableExtra)
kable(cars[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "latex", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)
```

# Inline code & results
Reproduction reaches new heights when you work with inline code. For instance, you can automatize the display of certain coefficients within the text. An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above. `r rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `r round(coef(model1)[2], 2)`. Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.      
Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects. For instance, imagine a reviewer asks you to omit certain observations from your sample. You can simply do so in the beginning of your code and push play subsequently.. at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.    
Researchers often avoid referring to results in-text etc. because you easily forget to change them when revising a manuscript. At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it. Inline code allows you to do just that. R Markdown allows you to that do so in a reproducible and automatized manner.

# Figures

## R base graphs
Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.rmd` file. It's relatively simple to insert R base graphs as you can see in Figure \@ref(fig:fig-1).

```{r fig-1, fig.align="center", fig.cap="Scatterplot of Speed and Distance", fig.pos="H", message=FALSE, warning=FALSE, paged.print=FALSE}
plot(cars$speed, cars$dist)
```

But it turns out that it doesn't always work so well.

## ggplot2 graphs
Same is true for ggplot2 as you can see in Figure \@ref(fig:fig-2).

```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
       shape="Number of \n Cylinders") + theme_classic()
```

# Good practices
Every researcher has his own optimized setup. Currently I would recommend the following:

* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders. You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).
* Make sure that filenames have a logic to them.
    + Main file with text/code: "paper.rmd", "report.rmd" 
    + Data files: "data_xxxxxx.*"
    + Image files: "fig_xxxxxx.*"
    + Tables files: "table_xxxx.*"
    + etc.
    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
* Name rchunks according to what they do or produce:
    + "`fig-...`" for chunks producing figures
    + "`table-...`" for chunks producing tables
    + "`model-...`" for chunks producing model estimates
    + "`import-...`" for chunks importing data
    + "`recoding-...`" for chunks in which data is recoded
* Ideally give chunks the labels of the figures/tables in the final publication, e.g., "`fig-1`" is the name of the chunk that produces Figure 1 in the final paper.
* Use "really" informative variable names:
    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher. 
    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
* Use unique identifiers in the final document:
    + e.g., name the models you estimate "M1", "M2" etc.
    + These unique names should also appear in the published paper.
    + Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code...

# Additional tricks for publishing
* Make your script anonymous
    + Simply put a `<!-- ... -->` around any identifying information, e.g., author names, title footnote etc.
* Counting words
    + Use adobe acrobat (commerical software) to convert your file to a word file. Then open in word and delete all the parts that shouldn't go into the word count. The word count is displayed in the lower right.
    + Use an one of the online services to count your words (search for "pdf word count")
* Appendix: You can change the numbering format for the appendix in the rmd file
    + What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
* Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file. See [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY) for a PLOS one example.

# Citation styles
If your study needs to follow a particular citation style, you can set the corresponding style in the header of your `.rmd` document. To do so you have to download the corresponding `.csl` file.

In the present document we use the style of the American Sociological Association and set it in the preamble with `csl: american-sociological-association.csl`. However, you also need to download the respective `.csl` file from the following github page: https://github.com/citation-style-language/styles and copy it into your working directory for it to work.

The github directory contains a wide variety of citation style files depending on what discipline you work in.

# References

\linespread{1}

::: {#refs}

:::

\clearpage

\appendix
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}

\renewcommand{\thesection}{A}

\setcounter{page}{1}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\figurename}{Table}

\setcounter{figure}{0}
\renewcommand\thefigure{A\arabic{figure}}
\renewcommand{\figurename}{Figure}

\clearpage
\pagenumbering{gobble}

\vspace*{7cm}

\begin{center}
\begin{huge}
Online Appendix
\end{huge}
\end{center}
\vspace{3cm}

\clearpage
\pagenumbering{arabic}

# Online appendix

## Attach R session info in appendix {#sec:rsessioninfo}

Since R and R packages are constantly evolving you might want to add the R session info that contains information on the R version as well as the packages that are loaded.

```{r echo=FALSE}
print(sessionInfo(), local = FALSE)
```

## All the code in the paper

To simply attach all the code you used in the PDF file in the appendix see the R chunk in the underlying `.rmd` file:

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

### 2021-06-16 paper.tex
Old files.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Times New Roman}
  \setsansfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{\vspace{1cm}Writing a reproducible paper\\
in Pagedown*\footnote{*Corresponding address: \href{mailto:mail@paulcbauer.eu}{\nolinkurl{mail@paulcbauer.eu}}. Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.}\vspace{0.5cm}\\}
\author{Paul C. Bauer \& Camille Landesvatter\\
Mannheim Centre for European Social Research}
\date{Based on: \url{https://osf.io/q395s}\\
This version: June 16, 2021\\
~\\
Feedback? Errors?\\
\href{mailto:mail@paulcbauer.eu}{\nolinkurl{mail@paulcbauer.eu}}\\
\href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{Github repository}}

\begin{document}
\maketitle
\begin{abstract}
\noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the ``tricks''/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded \href{https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing}{here}. I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{github repo}).\vspace{.8cm}
\end{abstract}

\setstretch{1.2}
\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

\hypertarget{why-reproducible-research-in-r}{%
\section{Why reproducible research (in R)?}\label{why-reproducible-research-in-r}}

Some arguments\ldots{}

\begin{itemize}
\tightlist
\item
  \textbf{Access}: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge (Kirsop and Chan \protect\hyperlink{ref-Kirsop2005-ro}{2005}).
\item
  \textbf{Reproducability}: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.
\item
  \textbf{Errors}: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to \textbf{automatize} such steps and/or avoid them.
\item
  \textbf{Revisions}: Revising a paper takes much less time if you have all the code you need in one place, i.e., one \texttt{.rmd} file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.
\end{itemize}

\hypertarget{prerequesites}{%
\section{Prerequesites}\label{prerequesites}}

I assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown. If you don't know what R Markdown is watch \href{https://vimeo.com/178485416}{this short video}. Then\ldots{}

\begin{itemize}
\tightlist
\item
  \ldots install \href{https://www.r-project.org/}{R} and \href{https://www.rstudio.com/}{Rstudio} (most recent versions) (R Core Team \protect\hyperlink{ref-R2017}{2017}; RStudio Team \protect\hyperlink{ref-Rstudio2015}{2015}).
\item
  \ldots install \href{https://yihui.name/tinytex/}{tinytex}, a lightweight version of Tex Live (Allaire et al. \protect\hyperlink{ref-markdown2017}{2017}; Xie \protect\hyperlink{ref-tinytex}{2018}\protect\hyperlink{ref-tinytex}{b}).
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{'tinytex'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{))}
\NormalTok{tinytex}\OperatorTok{::}\KeywordTok{install_tinytex}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  \ldots install the packages below using the code below (Sievert et al. \protect\hyperlink{ref-plotly}{2017}; Xie \protect\hyperlink{ref-knitr3}{2014}, \protect\hyperlink{ref-knitr2}{2015}, \protect\hyperlink{ref-bookdown2}{2016}, \protect\hyperlink{ref-bookdown1}{2017}, \protect\hyperlink{ref-knitr1}{2018}\protect\hyperlink{ref-knitr1}{a}; Zhu \protect\hyperlink{ref-kableextra}{2017}).
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"rmarkdown"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{,}
                   \StringTok{"stargazer"}\NormalTok{, }\StringTok{"plotly"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{,}
                   \StringTok{"bookdown"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\item
  \ldots download the 5 input files I created --- \texttt{paper.rmd}, \texttt{references.bib}, \texttt{data.csv} and \texttt{american-sociological-association.csl} --- from \href{https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing}{this folder}. Ignore the other files.
\item
  \ldots learn R and read about the other underlying components namely \href{https://en.wikipedia.org/wiki/Markdown}{Markdown}, \href{https://rmarkdown.rstudio.com/lesson-1.html}{R Markdown} and \href{https://en.wikipedia.org/wiki/LaTeX}{Latex}.
\end{itemize}

\hypertarget{basics-input-and-output-files}{%
\section{Basics: Input and output files}\label{basics-input-and-output-files}}

All the files you need to produce the present PDF file are the input files\ldots{}

\begin{itemize}
\tightlist
\item
  \ldots a \texttt{paper.rmd} file (the underlying R Markdown file).
\item
  \ldots a \texttt{references.bib} file (the bibliography).

  \begin{itemize}
  \tightlist
  \item
    I use paperpile to manage my references and export the \texttt{.bib} file into the folder that contains my \texttt{.rmd} file.
  \end{itemize}
\item
  \ldots a \texttt{data.csv} file (some raw data).
\item
  \ldots{} a \texttt{american-sociological-association.csl} file that defines the style of your bibliography.\footnote{You can download various citation style files from this webpage: \url{https://github.com/citation-style-language/styles}.}
\end{itemize}

\href{https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing}{Download these files} and save them into a folder. Close R/Rstudio and directly open \texttt{paper.rmd} with RStudio. Doing so assures that the working directory is set to the folder that contains \texttt{paper.rmd} and the other files.\footnote{You can always check your working directory in R with \texttt{getwd()}.}

Once you run/compile the \texttt{paper.rmd} file in Rstudio it creates mainly two output files:

\begin{itemize}
\tightlist
\item
  \texttt{paper.tex}
\item
  \texttt{paper.pdf} (the one you are reading right now)
\end{itemize}

In addition, there may be files that you generate and store locally in the folder during the compilation process. This is the case for some of the Plotly graphs below.

Ideally, we can simply provide others with a \texttt{zip} folder that contains both our input files and our output files. Then it's possible to reproduce the process from managing/analyzing some raw data to producing the final scientific article.

Below we always display the R code in the chunks that produce the output. In your paper you will normally only present outputs (e.g., tables, figures etc.) by choosing the chunk option ``Show output only'' in R Studio. The chunk commands itself are not displayed but they do matter for referencing etc. So simply orient yourself at the underlying \texttt{paper.rmd} file.

\hypertarget{referencing-within-your-document}{%
\section{Referencing within your document}\label{referencing-within-your-document}}

To see how referencing works simply see the different examples for figures, tables and sections below. For instance in Section \ref{sec:tables} you can find different ways of referencing tables. The code of the underlying \texttt{paper.rmd} will show you how I referenced Section \ref{sec:tables} right here namely with `\texttt{Section\ \textbackslash{}@ref(sec:tables)}'.

\hypertarget{software-versioning}{%
\section{Software versioning}\label{software-versioning}}

Software changes and gets updated, especially with an active developer community like that of R. Luckily you can always access \href{https://cran.r-project.org/bin/windows/base/old/}{old versions of R} and old version of R packages in \href{https://cran.r-project.org/src/contrib/Archive/}{the archive}. In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., \texttt{dplyr\_0.2.tar.gz}. Then insert the path in the following function: \texttt{install.packages("https://....../dplyr\_0.2.tar.gz",\ repos=NULL,\ type="source")}. Ideally, however, results will be simply reproducible in the most current R and package versions.

I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \ref{sec:rsessioninfo}. This will make sure you always provide the package versions that you used in the last compilation of your paper. For more advanced tools see \href{https://rstudio.github.io/packrat/}{packrat}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{cat}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"#"}\NormalTok{, }\KeywordTok{capture.output}\NormalTok{(}\KeywordTok{sessionInfo}\NormalTok{()), }\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{collapse =}\StringTok{""}\NormalTok{)) }
  \CommentTok{# or use message() instead of cat()}
\end{Highlighting}
\end{Shaded}

\hypertarget{data}{%
\section{Data}\label{data}}

\hypertarget{import}{%
\subsection{Import}\label{import}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{data <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data.csv"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   X speed dist
## 1 1     4    2
## 2 2     4   10
## 3 3     7    4
## 4 4     7   22
## 5 5     8   16
## 6 6     9   10
\end{verbatim}

\hypertarget{putting-your-entire-data-into-the-.rmd-file}{%
\subsection{Putting your entire data into the .rmd file}\label{putting-your-entire-data-into-the-.rmd-file}}

Applying the function \texttt{dput()} to an object gives you the code needed to reproduce that object. So you could paste that code into your \texttt{.rmd} file if you don't want to have extra data files. This makes sense were data files are small.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dput}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## structure(list(X = 1:50, speed = c(4L, 4L, 7L, 7L, 8L, 9L, 10L, 
## 10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, 
## 14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, 
## 18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, 
## 24L, 24L, 24L, 25L), dist = c(2L, 10L, 4L, 22L, 16L, 10L, 18L, 
## 26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, 
## 36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, 
## 76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, 
## 92L, 93L, 120L, 85L)), class = "data.frame", row.names = c(NA, 
## -50L))
\end{verbatim}

You can then insert the dput output in your \texttt{.rmd} as below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{data <-}\StringTok{ }\KeywordTok{structure}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{X =} \DecValTok{1}\OperatorTok{:}\DecValTok{50}\NormalTok{, }\DataTypeTok{speed =} \KeywordTok{c}\NormalTok{(4L, 4L, 7L, 7L, 8L, 9L, 10L, }
\NormalTok{10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, }
\NormalTok{14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, }
\NormalTok{18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, }
\NormalTok{24L, 24L, 24L, 25L), }\DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(2L, 10L, 4L, 22L, 16L, 10L, 18L, }
\NormalTok{26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, }
\NormalTok{36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, }
\NormalTok{76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, }
\NormalTok{92L, 93L, 120L, 85L)), }
\DataTypeTok{class =} \StringTok{"data.frame"}\NormalTok{, }\DataTypeTok{row.names =} \KeywordTok{c}\NormalTok{(}\OtherTok{NA}\NormalTok{, }
\OperatorTok{-}\NormalTok{50L))}
\end{Highlighting}
\end{Shaded}

\hypertarget{sec:tables}{%
\section{Tables}\label{sec:tables}}

Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for \texttt{stargazer} the label is contained in the function, for \texttt{kable} it's contained in the chunk name.

\hypertarget{stargazer-summary-and-regression-tables}{%
\subsection{stargazer(): Summary and regression tables}\label{stargazer-summary-and-regression-tables}}

Table \ref{tab1} shows summary stats of your data.\footnote{To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.} I normally use \texttt{stargazer()} (Hlavac \protect\hyperlink{ref-hlavac2013stargazer}{2013}) which offers extreme flexibility regarding table output (see \texttt{?stargazer}).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stargazer)}
\KeywordTok{stargazer}\NormalTok{(cars, }
          \DataTypeTok{title =} \StringTok{"Summary table with stargazer"}\NormalTok{,}
          \DataTypeTok{label=}\StringTok{"tab1"}\NormalTok{, }
          \DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{, }
          \DataTypeTok{header=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[H] \centering 
  \caption{Summary table with stargazer} 
  \label{tab1} 
\begin{tabular}{@{\extracolsep{5pt}}lccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Pctl(25)} & \multicolumn{1}{c}{Pctl(75)} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
speed & 50 & 15.400 & 5.288 & 4 & 12 & 19 & 25 \\ 
dist & 50 & 42.980 & 25.769 & 2 & 26 & 56 & 120 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

Table \ref{tab2} shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stargazer)}
\NormalTok{model1 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(speed }\OperatorTok{~}\StringTok{ }\NormalTok{dist, }\DataTypeTok{data =}\NormalTok{ cars)}
\NormalTok{model2 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(speed }\OperatorTok{~}\StringTok{ }\NormalTok{dist, }\DataTypeTok{data =}\NormalTok{ cars)}
\NormalTok{model3 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(dist }\OperatorTok{~}\StringTok{ }\NormalTok{speed, }\DataTypeTok{data =}\NormalTok{ cars)}
\KeywordTok{stargazer}\NormalTok{(model1, model2, model3,}
          \DataTypeTok{title =} \StringTok{"Regression table with stargazer"}\NormalTok{,}
          \DataTypeTok{label=}\StringTok{"tab2"}\NormalTok{, }
          \DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{, }
          \DataTypeTok{column.labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"M1"}\NormalTok{, }\StringTok{"M2"}\NormalTok{, }\StringTok{"M3"}\NormalTok{),}
          \DataTypeTok{model.numbers =} \OtherTok{FALSE}\NormalTok{,}
          \DataTypeTok{header=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[H] \centering 
  \caption{Regression table with stargazer} 
  \label{tab2} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & \multicolumn{2}{c}{speed} & dist \\ 
 & M1 & M2 & M3 \\ 
\hline \\[-1.8ex] 
 dist & 0.166$^{***}$ & 0.166$^{***}$ &  \\ 
  & (0.017) & (0.017) &  \\ 
  & & & \\ 
 speed &  &  & 3.932$^{***}$ \\ 
  &  &  & (0.416) \\ 
  & & & \\ 
 Constant & 8.284$^{***}$ & 8.284$^{***}$ & $-$17.579$^{**}$ \\ 
  & (0.874) & (0.874) & (6.758) \\ 
  & & & \\ 
\hline \\[-1.8ex] 
Observations & 50 & 50 & 50 \\ 
R$^{2}$ & 0.651 & 0.651 & 0.651 \\ 
Adjusted R$^{2}$ & 0.644 & 0.644 & 0.644 \\ 
Residual Std. Error (df = 48) & 3.156 & 3.156 & 15.380 \\ 
F Statistic (df = 1; 48) & 89.567$^{***}$ & 89.567$^{***}$ & 89.567$^{***}$ \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\hypertarget{kable-and-kable_styling}{%
\subsection{kable() and kable\_styling()}\label{kable-and-kable_styling}}

Another great function is \texttt{kable()} (\texttt{knitr} package) in combination with \texttt{kableExtra}. Table \ref{tab:tab3} provides an example.\footnote{To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.} Again you can modify so many things in both the \texttt{kable()} and the \texttt{kable\_styling()} function. See \href{https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf}{this overview} of all the kable stylings that are possible provided by the package author himself.

\begin{table}[H]

\caption{\label{tab:tab3}Table with kable() and kablestyling()}
\centering
\resizebox{\linewidth}{!}{
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X}
\toprule
  & speed & dist\\
\midrule
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{2}\\
2 & 4 & 10\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{4}\\
4 & 7 & 22\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{8} & \cellcolor{gray!6}{16}\\
\addlinespace
6 & 9 & 10\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{10} & \cellcolor{gray!6}{18}\\
8 & 10 & 26\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{10} & \cellcolor{gray!6}{34}\\
10 & 11 & 17\\
\bottomrule
\end{tabu}}
\end{table}

\hypertarget{inline-code-results}{%
\section{Inline code \& results}\label{inline-code-results}}

Reproduction reaches new heights when you work with inline code. For instance, you can automatize the display of certain coefficients within the text. An example is to include estimates, e.g., the coefficient of \texttt{dist} of the model we ran above. \texttt{\textasciigrave{}r\ round(coef(model1){[}2{]},\ 2)\textasciigrave{}} will insert the coefficient as follows: 0.17. Or \texttt{\textasciigrave{}r\ 3\ +\ 7\textasciigrave{}} will insert a 10 in the text.\\
Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects. For instance, imagine a reviewer asks you to omit certain observations from your sample. You can simply do so in the beginning of your code and push play subsequently.. at time you might have to set \texttt{cache\ =\ FALSE} at the beginning so that all the code chunks are rerun.\\
Researchers often avoid referring to results in-text etc. because you easily forget to change them when revising a manuscript. At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it. Inline code allows you to do just that. R Markdown allows you to that do so in a reproducible and automatized manner.

\hypertarget{figures}{%
\section{Figures}\label{figures}}

\hypertarget{r-base-graphs}{%
\subsection{R base graphs}\label{r-base-graphs}}

Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the \texttt{.rmd} file. It's relatively simple to insert R base graphs as you can see in Figure \ref{fig:fig-1}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(cars}\OperatorTok{$}\NormalTok{speed, cars}\OperatorTok{$}\NormalTok{dist)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-latex/fig-1-1}

}

\caption{Scatterplot of Speed and Distance}\label{fig:fig-1}
\end{figure}

But it turns out that it doesn't always work so well.

\hypertarget{ggplot2-graphs}{%
\subsection{ggplot2 graphs}\label{ggplot2-graphs}}

Same is true for ggplot2 as you can see in Figure \ref{fig:fig-2}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mtcars}\OperatorTok{$}\NormalTok{cyl <-}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(mtcars}\OperatorTok{$}\NormalTok{cyl) }\CommentTok{# Convert cyl to factor}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{ggplot}\NormalTok{(mtcars, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{wt, }\DataTypeTok{y=}\NormalTok{mpg, }\DataTypeTok{shape=}\NormalTok{cyl)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{"Weight (lb/1000)"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Miles/(US) gallon"}\NormalTok{, }
       \DataTypeTok{shape=}\StringTok{"Number of }\CharTok{\textbackslash{}n}\StringTok{ Cylinders"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-latex/fig-2-1}

}

\caption{Miles per gallon according to the weight}\label{fig:fig-2}
\end{figure}

\hypertarget{good-practices}{%
\section{Good practices}\label{good-practices}}

Every researcher has his own optimized setup. Currently I would recommend the following:

\begin{itemize}
\tightlist
\item
  Keep all files of your project (that matter for producing the PDF) in one folder without subfolders. You can zip and directly upload that folder to the \href{https://dataverse.harvard.edu/}{Harvard dataverse}.
\item
  Make sure that filenames have a logic to them.

  \begin{itemize}
  \tightlist
  \item
    Main file with text/code: ``paper.rmd'', ``report.rmd''
  \item
    Data files: "data\_xxxxxx.*"
  \item
    Image files: "fig\_xxxxxx.*"
  \item
    Tables files: "table\_xxxx.*"
  \item
    etc.
  \item
    Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called \texttt{fig\_1\_xxxxx.pdf}.
  \end{itemize}
\item
  Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
\item
  Name rchunks according to what they do or produce:

  \begin{itemize}
  \tightlist
  \item
    ``\texttt{fig-...}'' for chunks producing figures
  \item
    ``\texttt{table-...}'' for chunks producing tables
  \item
    ``\texttt{model-...}'' for chunks producing model estimates
  \item
    ``\texttt{import-...}'' for chunks importing data
  \item
    ``\texttt{recoding-...}'' for chunks in which data is recoded
  \end{itemize}
\item
  Ideally give chunks the labels of the figures/tables in the final publication, e.g., ``\texttt{fig-1}'' is the name of the chunk that produces Figure 1 in the final paper.
\item
  Use ``really'' informative variable names:

  \begin{itemize}
  \tightlist
  \item
    Q: What do you think does the variable \emph{trstep} measure? It actually measures trust in the European parliament.

    \begin{itemize}
    \tightlist
    \item
      How could we call this variable instead? Yes, \texttt{trust.european.parliament} which is longer but will probably be understood by another researcher.
    \end{itemize}
  \item
    If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
  \end{itemize}
\item
  Use unique identifiers in the final document:

  \begin{itemize}
  \tightlist
  \item
    e.g., name the models you estimate ``M1'', ``M2'' etc.
  \item
    These unique names should also appear in the published paper.
  \item
    Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code\ldots{}
  \end{itemize}
\end{itemize}

\hypertarget{additional-tricks-for-publishing}{%
\section{Additional tricks for publishing}\label{additional-tricks-for-publishing}}

\begin{itemize}
\tightlist
\item
  Make your script anonymous

  \begin{itemize}
  \tightlist
  \item
    Simply put a \texttt{\textless{}!-\/-\ ...\ -\/-\textgreater{}} around any identifying information, e.g., author names, title footnote etc.
  \end{itemize}
\item
  Counting words

  \begin{itemize}
  \tightlist
  \item
    Use adobe acrobat (commerical software) to convert your file to a word file. Then open in word and delete all the parts that shouldn't go into the word count. The word count is displayed in the lower right.
  \item
    Use an one of the online services to count your words (search for ``pdf word count'')
  \end{itemize}
\item
  Appendix: You can change the numbering format for the appendix in the rmd file

  \begin{itemize}
  \tightlist
  \item
    What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
  \end{itemize}
\item
  Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file. See \href{https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY}{here} for a PLOS one example.
\end{itemize}

\hypertarget{citation-styles}{%
\section{Citation styles}\label{citation-styles}}

If your study needs to follow a particular citation style, you can set the corresponding style in the header of your \texttt{.rmd} document. To do so you have to download the corresponding \texttt{.csl} file.

In the present document we use the style of the American Sociological Association and set it in the preamble with \texttt{csl:\ american-sociological-association.csl}. However, you also need to download the respective \texttt{.csl} file from the following github page: \url{https://github.com/citation-style-language/styles} and copy it into your working directory for it to work.

The github directory contains a wide variety of citation style files depending on what discipline you work in.

\hypertarget{references}{%
\section{References}\label{references}}

\linespread{1}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-markdown2017}{}%
Allaire, JJ, Jeffrey Horner, Vicent Marti, and Natacha Porte. 2017. \emph{Markdown: 'Markdown' Rendering for R}.

\leavevmode\hypertarget{ref-hlavac2013stargazer}{}%
Hlavac, Marek. 2013. ``Stargazer: LaTeX Code and Ascii Text for Well-Formatted Regression and Summary Statistics Tables.'' \emph{URL: Http://CRAN. R-Project. Org/Package= Stargazer}.

\leavevmode\hypertarget{ref-Kirsop2005-ro}{}%
Kirsop, Barbara and Leslie Chan. 2005. ``Transforming Access to Research Literature for Developing Countries.'' \emph{Serials Review} 31(4):246--55.

\leavevmode\hypertarget{ref-R2017}{}%
R Core Team. 2017. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Rstudio2015}{}%
RStudio Team. 2015. \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc.

\leavevmode\hypertarget{ref-plotly}{}%
Sievert, Carson, Chris Parmer, Toby Hocking, Scott Chamberlain, Karthik Ram, Marianne Corvellec, and Pedro Despouy. 2017. \emph{Plotly: Create Interactive Web Graphics via 'Plotly.js'}.

\leavevmode\hypertarget{ref-knitr3}{}%
Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible Research in R.'' in \emph{Implementing reproducible computational research}, edited by V. Stodden, F. Leisch, and R. D. Peng. Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-knitr2}{}%
Xie, Yihui. 2015. \emph{Dynamic Documents with R and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-bookdown2}{}%
Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-bookdown1}{}%
Xie, Yihui. 2017. \emph{Bookdown: Authoring Books and Technical Documents with R Markdown}.

\leavevmode\hypertarget{ref-knitr1}{}%
Xie, Yihui. 2018a. \emph{Knitr: A General-Purpose Package for Dynamic Report Generation in R}.

\leavevmode\hypertarget{ref-tinytex}{}%
Xie, Yihui. 2018b. \emph{Tinytex: Helper Functions to Install and Maintain 'Tex Live', and Compile 'Latex' Documents}.

\leavevmode\hypertarget{ref-kableextra}{}%
Zhu, Hao. 2017. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}.

\clearpage

\appendix
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}

\renewcommand{\thesection}{A}

\setcounter{page}{1}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\figurename}{Table}

\setcounter{figure}{0}
\renewcommand\thefigure{A\arabic{figure}}
\renewcommand{\figurename}{Figure}

\clearpage
\pagenumbering{gobble}

\vspace*{7cm}

\begin{center}
\begin{huge}
Online Appendix
\end{huge}
\end{center}
\vspace{3cm}

\clearpage
\pagenumbering{arabic}

\hypertarget{online-appendix}{%
\section{Online appendix}\label{online-appendix}}

\hypertarget{sec:rsessioninfo}{%
\subsection{Attach R session info in appendix}\label{sec:rsessioninfo}}

Since R and R packages are constantly evolving you might want to add the R session info that contains information on the R version as well as the packages that are loaded.

\begin{verbatim}
## R version 4.1.0 (2021-05-18)
## Platform: x86_64-w64-mingw32/x64 (64-bit)
## Running under: Windows 10 x64 (build 17763)
## 
## Matrix products: default
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] ggplot2_3.3.4    kableExtra_1.3.4 knitr_1.33       stargazer_5.2.2 
## 
## loaded via a namespace (and not attached):
##  [1] compiler_4.1.0    pillar_1.6.1      tools_4.1.0       digest_0.6.27    
##  [5] evaluate_0.14     lifecycle_1.0.0   tibble_3.1.2      gtable_0.3.0     
##  [9] viridisLite_0.4.0 pkgconfig_2.0.3   rlang_0.4.11      DBI_1.1.1        
## [13] rstudioapi_0.13   yaml_2.2.1        xfun_0.24         withr_2.4.2      
## [17] dplyr_1.0.6       httr_1.4.2        stringr_1.4.0     xml2_1.3.2       
## [21] generics_0.1.0    vctrs_0.3.8       systemfonts_1.0.2 tidyselect_1.1.1 
## [25] webshot_0.5.2     grid_4.1.0        svglite_2.0.0     glue_1.4.2       
## [29] R6_2.5.0          fansi_0.5.0       rmarkdown_2.9     bookdown_0.22    
## [33] farver_2.1.0      purrr_0.3.4       magrittr_2.0.1    scales_1.1.1     
## [37] htmltools_0.5.1.1 ellipsis_0.3.2    assertthat_0.2.1  rvest_1.0.0      
## [41] colorspace_2.0-1  labeling_0.4.2    utf8_1.2.1        stringi_1.6.1    
## [45] munsell_0.5.0     crayon_1.4.1
\end{verbatim}

\hypertarget{all-the-code-in-the-paper}{%
\subsection{All the code in the paper}\label{all-the-code-in-the-paper}}

To simply attach all the code you used in the PDF file in the appendix see the R chunk in the underlying \texttt{.rmd} file:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\OperatorTok{::}\NormalTok{opts_chunk}\OperatorTok{$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{cache =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# Use chache = TRUE if you want to speed up compilation}

\CommentTok{# A function to allow for showing some of the inline code}
\NormalTok{rinline <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(code)\{}
\NormalTok{  html <-}\StringTok{ '<code  class="r">``` `r CODE` ```</code>'}
  \KeywordTok{sub}\NormalTok{(}\StringTok{"CODE"}\NormalTok{, code, html)}
\NormalTok{\}}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{'tinytex'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{))}
\NormalTok{tinytex}\OperatorTok{::}\KeywordTok{install_tinytex}\NormalTok{()}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"rmarkdown"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{,}
                   \StringTok{"stargazer"}\NormalTok{, }\StringTok{"plotly"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{,}
                   \StringTok{"bookdown"}\NormalTok{))}
\KeywordTok{cat}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"#"}\NormalTok{, }\KeywordTok{capture.output}\NormalTok{(}\KeywordTok{sessionInfo}\NormalTok{()), }\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{collapse =}\StringTok{""}\NormalTok{)) }
  \CommentTok{# or use message() instead of cat()}
\NormalTok{data <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data.csv"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(data)}
\KeywordTok{dput}\NormalTok{(data)}
\NormalTok{data <-}\StringTok{ }\KeywordTok{structure}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{X =} \DecValTok{1}\OperatorTok{:}\DecValTok{50}\NormalTok{, }\DataTypeTok{speed =} \KeywordTok{c}\NormalTok{(4L, 4L, 7L, 7L, 8L, 9L, 10L, }
\NormalTok{10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, }
\NormalTok{14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, }
\NormalTok{18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, }
\NormalTok{24L, 24L, 24L, 25L), }\DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(2L, 10L, 4L, 22L, 16L, 10L, 18L, }
\NormalTok{26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, }
\NormalTok{36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, }
\NormalTok{76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, }
\NormalTok{92L, 93L, 120L, 85L)), }
\DataTypeTok{class =} \StringTok{"data.frame"}\NormalTok{, }\DataTypeTok{row.names =} \KeywordTok{c}\NormalTok{(}\OtherTok{NA}\NormalTok{, }
\OperatorTok{-}\NormalTok{50L))}
\KeywordTok{library}\NormalTok{(stargazer)}
\KeywordTok{stargazer}\NormalTok{(cars, }
          \DataTypeTok{title =} \StringTok{"Summary table with stargazer"}\NormalTok{,}
          \DataTypeTok{label=}\StringTok{"tab1"}\NormalTok{, }
          \DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{, }
          \DataTypeTok{header=}\OtherTok{FALSE}\NormalTok{)}
\KeywordTok{library}\NormalTok{(stargazer)}
\NormalTok{model1 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(speed }\OperatorTok{~}\StringTok{ }\NormalTok{dist, }\DataTypeTok{data =}\NormalTok{ cars)}
\NormalTok{model2 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(speed }\OperatorTok{~}\StringTok{ }\NormalTok{dist, }\DataTypeTok{data =}\NormalTok{ cars)}
\NormalTok{model3 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(dist }\OperatorTok{~}\StringTok{ }\NormalTok{speed, }\DataTypeTok{data =}\NormalTok{ cars)}
\KeywordTok{stargazer}\NormalTok{(model1, model2, model3,}
          \DataTypeTok{title =} \StringTok{"Regression table with stargazer"}\NormalTok{,}
          \DataTypeTok{label=}\StringTok{"tab2"}\NormalTok{, }
          \DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{, }
          \DataTypeTok{column.labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"M1"}\NormalTok{, }\StringTok{"M2"}\NormalTok{, }\StringTok{"M3"}\NormalTok{),}
          \DataTypeTok{model.numbers =} \OtherTok{FALSE}\NormalTok{,}
          \DataTypeTok{header=}\OtherTok{FALSE}\NormalTok{)}
\KeywordTok{library}\NormalTok{(knitr)}
\KeywordTok{library}\NormalTok{(kableExtra)}
\KeywordTok{kable}\NormalTok{(cars[}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{,], }\DataTypeTok{row.names =} \OtherTok{TRUE}\NormalTok{, }
      \DataTypeTok{caption =} \StringTok{'Table with kable() and kablestyling()'}\NormalTok{, }
      \DataTypeTok{format =} \StringTok{"latex"}\NormalTok{, }\DataTypeTok{booktabs =}\NormalTok{ T) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{full_width =}\NormalTok{ T, }
                      \DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }
                                        \StringTok{"scale_down"}\NormalTok{,}
                                        \StringTok{"HOLD_position"}\NormalTok{),}
                      \DataTypeTok{font_size =} \DecValTok{10}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(cars}\OperatorTok{$}\NormalTok{speed, cars}\OperatorTok{$}\NormalTok{dist)}
\NormalTok{mtcars}\OperatorTok{$}\NormalTok{cyl <-}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(mtcars}\OperatorTok{$}\NormalTok{cyl) }\CommentTok{# Convert cyl to factor}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{ggplot}\NormalTok{(mtcars, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{wt, }\DataTypeTok{y=}\NormalTok{mpg, }\DataTypeTok{shape=}\NormalTok{cyl)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{"Weight (lb/1000)"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Miles/(US) gallon"}\NormalTok{, }
       \DataTypeTok{shape=}\StringTok{"Number of }\CharTok{\textbackslash{}n}\StringTok{ Cylinders"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_classic}\NormalTok{()}
\KeywordTok{print}\NormalTok{(}\KeywordTok{sessionInfo}\NormalTok{(), }\DataTypeTok{local =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2021-06-17 paper_p.Rmd
First commit CSS files and added paper_p.

---
title: "Writing a reproducible paper in Pagedown^[Corresponding address: mail@paulcbauer.eu.<br> Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]"
shorttitle: "**foo**:  A Capitalized Title"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <landesvatterc@gmail.com>
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded [here]<https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
# at least one keyword must be supplied
keywords: [keywords, not capitalized, "[Java]{.proglang}"]
paged-footnotes: true
link-citations: true
linestretch: 1.2
colorlinks: true
bibliography: index.bib
output:
  pagedown::html_paged:
    toc: true
    toc_depth: 3    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: TRUE
    css: ['default.css', 'default-fonts.css', 'default-page.css']
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
---

# Introduction

This template demonstrates some of the basic [markdown]{.proglang} you'll need to know to create a JSS article with **pagedown** [@R-pagedown].

## Code formatting

You should use **Pandoc**'s [markdown]{.proglang} but you also can use $\LaTeX$ commands:

* [Java]{.proglang} or $\proglang{Java}$ 
* **plyr**, [plyr]{.pkg} or $\pkg{plyr}$
* `print("abc")` or $\code{print("abc")}$

# [R]{.proglang} code

Can be inserted in regular [R Markdown]{.proglang} blocks.

```{r}
x <- 1:10
x
```

# Render the document

If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.

# References {-}

```{r, include=FALSE}
knitr::write_bib(c(.packages(), 'pagedown', 'xaringan'), 'index.bib')
```

### 2021-06-18 paper_pagedown.Rmd
replaced default CSS (html_paged) wit jss CSS (jss_paged).

---
title: "Writing a reproducible paper in Pagedown^[Corresponding address: mail@paulcbauer.eu.<br> Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]"
shorttitle: "Writing a reproducible paper in Pagedown"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <landesvatterc@gmail.com>
date: |
  |       
  | Based on: https://osf.io/q395s
  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
  |
  | Feedback? Errors?
  | mail@paulcbauer.eu
  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
abstract: 'The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded [here](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).'
# at least one keyword must be supplied
keywords: [keywords]
links-to-footnotes: true
paged-footnotes: true
bibliography: index.bib
output: 
  pagedown::jss_paged:
    self_contained: false
    css: ['jss.css', 'jss-fonts.css', 'jss-page.css']
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
---

\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

\newpage
# Why reproducible research (in R)?

\newpage
# Prerequesites

\newpage
# Basics: Input and output files

\newpage
# Referencing within your document

\newpage
# Software versioning

\newpage
# Inline code & results

Can be inserted in regular [R Markdown]{.proglang} blocks.

```{r}
x <- 1:10
x
```

\newpage
# Data
## Import

```{r, echo=T, results="raw"}
data <- read.csv("data.csv")
head(data)
```

# Tables {#sec:tables}

Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.

## stargazer(): Summary and regression tables

Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).

```{r, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab1", 
          table.placement = "H", 
          type="html",
          header=FALSE)
```

Table \@ref(tab2) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(stargazer)
model1 <- lm(speed ~ dist, data = cars)
model2 <- lm(speed ~ dist, data = cars)
model3 <- lm(dist ~ speed, data = cars)

stargazer(model1, model2, model3,
          title = "Regression table with stargazer",
          label="tab2",
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          type="html",
          model.numbers = FALSE,
          header=FALSE)
```

## kable() and kable_styling()

Another great function is `kable()` (`knitr` package) in combination with `kableExtra`. Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.]

```{r tab3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(knitr)
library(kableExtra)

kable(mtcars[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)

```

# Render the document

If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.

# References {-}

```{r, include=FALSE}
knitr::write_bib(c(.packages(), 'pagedown', 'xaringan'), 'index.bib')
```

### 2021-06-20 paper_pagedown.Rmd
added a date class to the front page (wp_paged.html) and modified its size and appearance (jss.css).

@@
     email: <landesvatterc@gmail.com>
-date: |
-  |       
-  | Based on: https://osf.io/q395s
-  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
-  |
-  | Feedback? Errors?
-  | mail@paulcbauer.eu
-  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
+date: 'First version: 2 December, 2019<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
 abstract: 'The present paper provides a template for a reproducible scientific paper written in R Markdown.
@@
   pagedown::jss_paged:
+    template: wp_paged.html
     self_contained: false

### 2021-06-21 paper_pagedown.Rmd
modified rmd template.

@@
 ---
-title: "Writing a reproducible paper in Pagedown^[Corresponding address: mail@paulcbauer.eu.<br> Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]"
+title: "Writing a reproducible paper in Pagedown^[Corresponding adress: mail@paulcbauer.eu]"
 shorttitle: "Writing a reproducible paper in Pagedown"
@@
     affiliation: Mannheim Centre for European Social Research
-    email: <landesvatterc@gmail.com>
-date: 'First version: 2 December, 2019<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
+    email: <camille.landesvatter@mzes.uni-mannheim.de>
+date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
 abstract: 'The present paper provides a template for a reproducible scientific paper written in R Markdown.
@@
 # at least one keyword must be supplied
-keywords: [keywords]
+keywords: [open science, transparency, replication, R, markdown, pagedown]
 links-to-footnotes: true
@@
     template: wp_paged.html
-    self_contained: false
+    self_contained: true
     css: ['jss.css', 'jss-fonts.css', 'jss-page.css']
 # uncomment this line to produce HTML and PDF in RStudio:
-#knit: pagedown::chrome_print
+knit: pagedown::chrome_print
 ---
@@
 
+Some arguments...
+
+* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers).
+Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software.
+Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].
+* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
+A fully reproducible setup will help you to trace back your own steps.
+Obviously, the same is true for other researchers who may want to understand your work and built on it.
+It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.
+* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
+* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file.
+For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.
+
+# Why Pagedown?
+
 \newpage
 # Prerequesites
+
+We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown.
+If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416).
+Also, there is an overview and template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).
+The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.
+
+Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents.
+For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing R's pagedown given by Yihui Xie (who in addition to Romain Lesur developed R's pagedown package).
+If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).
+
+Then...
+
+* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
+* ...install the "pagedown"-package from github using the code below.
+```{r eval=FALSE, include=TRUE}
+remotes::install_github('rstudio/pagedown')
+```
+* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
+```{r eval=FALSE, include=TRUE}
+install.packages(c("rmarkdown", "knitr", "kableExtra",
+                   "stargazer", "plotly", "knitr"))
+```
+* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
+Ignore the other files.
+* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
+* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
+* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
+* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style.
+If however you want to modify single aspects or create your own template, you will need at least some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
 
@@
 
+All the files you need to produce the present PDF file are:
+
+1. the input files:
+
+* `paper.rmd` (the underlying R Markdown file).
+* `references.bib` (the bibliography).
+    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.rmd` file.
+* `data.csv` (some raw data).
+* `american-sociological-association.csl` (defines the style of your bibliography).^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]
+
+2. the "styling" files:
+
+Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document.
+With using our templates, you will create a document that has the "look" of a working paper.
+
+* `wp_paged.html`
+* `wp.css`
+* `wp-fonts.css`
+* `wp-pages.css`
+
+Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line 18-22) to see how to specifiy these files.
+Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom css and html.
+
+[Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder.
+Close R/Rstudio and directly open `paper.rmd` with RStudio.
+Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]
+
+Once you run/compile the `paper.rmd` file in Rstudio it creates mainly one output files:
+
+* `paper_pagedown.html` (the one you are reading right now)
+
+By using pagedown's `chrome_print` function in the YAML (uncomment the line) your html based web page will be printed to be PDF (the PDF will be stored in your working directory)
+
+* `paper_pagedown.pdf`
+
 \newpage
@@
 
-\newpage
+To see how referencing works simply see the different examples for figures, tables and sections below.
+For instance in Section \@ref(sec:tables) you can find different ways of referencing tables.
+The code of the underlying `paper.rmd` will show you how I referenced Section \@ref(sec:tables) right here namely with '`Section \@ref(sec:tables)`'.
+
 # Software versioning
 
+Software changes and gets updated, especially with an active developer community like that of R.
+Luckily you can always access [old versions of R](https://cran.r-project.org/bin/windows/base/old/) and old version of R packages in [the archive](https://cran.r-project.org/src/contrib/Archive/).
+In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., `dplyr_0.2.tar.gz`.
+Then insert the path in the following function: `install.packages("https://....../dplyr_0.2.tar.gz", repos=NULL, type="source")`.
+Ideally, however, results will be simply reproducible in the most current R and package versions.
+
+I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \@ref(sec:rsessioninfo).
+This will make sure you always provide the package versions that you used in the last compilation of your paper.
+For more advanced tools see [packrat](https://rstudio.github.io/packrat/).
+
+```{r fig-versioning, echo=TRUE, eval=FALSE}
+cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
+  # or use message() instead of cat()
+```
+
 \newpage
-# Inline code & results
+# Data
+## Import
 
-Can be inserted in regular [R Markdown]{.proglang} blocks.
+Generally, code is evaluated by inserting regular [R Markdown]{.proglang} blocks.
 
@@
 
-\newpage
-# Data
-## Import
+Below we import an exemplary dataset ([download](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing)).
 
@@
 
+### Summary table with stargazer
+
 Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
@@
 stargazer(cars, 
-          title = "Summary table with stargazer",
           label="tab1", 
@@
 ```
+
+### Regression table with stargazer
 
@@
 stargazer(model1, model2, model3,
-          title = "Regression table with stargazer",
           label="tab2",
@@
 
-# Render the document
+## modelsummary
 
-If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio.
+https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html
+
+# Inline code & results
+
+# Figures
+
+## R base graphs
+Inserting figures can be slightly more complicated.
+Ideally, we would produce and insert them directly in the `.rmd` file.
+It's relatively simple to insert R base graphs as you can see in Figure \@ref(fig:fig-1).
+
+```{r fig-1, fig.align="center", fig.cap="Scatterplot of Speed and Distance", fig.pos="H", message=FALSE, warning=FALSE, paged.print=FALSE}
+plot(cars$speed, cars$dist)
+```
+
+But it turns out that it doesn't always work so well.
+
+## ggplot2 graphs
+Same is true for ggplot2 as you can see in Figure \@ref(fig:fig-2).
+
+```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
+mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
+library(ggplot2)
+ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
+  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
+       shape="Number of \n Cylinders") + theme_classic()
+```
+
+## Interactive graphs
+
+html!
+
+# Compiling
+
+To view your paper, pagedown requires a web server (since it is based von paged.js)^(open-source library to paginate content in the browser).
+By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.
+
+There are several options, depending on your intention.
+
+- click on the `Knit` button which by default will provide a html document in the RStudio viewer pane (the html will be stored in your working directory)
+- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if you want your html based web page to be (additionally to the html!) printed to be PDF (the PDF will be stored in your working directory)
+- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.
+You can simply call the function `xaringan::inf_mr()` (within your console).
+This will launch a local web server via the **servr** package (Xie 2021a) and display your pages in the RStudio viewer.
+Each time you save your document (_Ctrl+S_) xaringan updates your pages in the viewer.
+
+- If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio.
 Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.
+
+# Good practices
+Every researcher has his own optimized setup. Currently I would recommend the following:
+
+* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders.
+You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).
+* Make sure that filenames have a logic to them.
+    + Main file with text/code: "paper.rmd", "report.rmd" 
+    + Data files: "data_xxxxxx.*"
+    + Image files: "fig_xxxxxx.*"
+    + Tables files: "table_xxxx.*"
+    + etc.
+    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
+* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
+* Name rchunks according to what they do or produce:
+    + "`fig-...`" for chunks producing figures
+    + "`table-...`" for chunks producing tables
+    + "`model-...`" for chunks producing model estimates
+    + "`import-...`" for chunks importing data
+    + "`recoding-...`" for chunks in which data is recoded
+* Use "really" informative variable names:
+    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
+        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher. 
+    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
+* Use unique identifiers in the final document:
+    + e.g., name the models you estimate "M1", "M2" etc.
+    + These unique names should also appear in the published paper.
+    + Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code...
+
+# Additional tricks for publishing
+* Make your script anonymous
+    + Simply put a `<!-- ... -->` around any identifying information, e.g., author names, title footnote etc.
+* Counting words
+    + Use adobe acrobat (commerical software) to convert your file to a word file.
+Then open in word and delete all the parts that shouldn't go into the word count.
+The word count is displayed in the lower right.
+    + Use an one of the online services to count your words (search for "pdf word count")
+* Appendix: You can change the numbering format for the appendix in the rmd file
+    + What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
+* Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file.
+See [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY) for a PLOS one example.
+
+# Citation styles
+If your study needs to follow a particular citation style, you can set the corresponding style in the header of your `.rmd` document. To do so you have to download the corresponding `.csl` file.
+
+In the present document we use the style of the American Sociological Association and set it in the preamble with `csl: american-sociological-association.csl`.
+However, you also need to download the respective `.csl` file from the following github page: https://github.com/citation-style-language/styles and copy it into your working directory for it to work.
+
+The github directory contains a wide variety of citation style files depending on what discipline you work in.
 
@@
 ```
+
+# Online appendix
+
+## Attach R session info in appendix {#sec:rsessioninfo}
+
+Since R and R packages are constantly evolving you might want to add the R session info that contains information on the R version as well as the packages that are loaded.
+
+```{r echo=FALSE}
+print(sessionInfo(), local = FALSE)
+```
+
+## All the code in the paper
+
+To simply attach all the code you used in the PDF file in the appendix see the R chunk in the underlying `.rmd` file:
+
+```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
+```

### 2021-06-21 paper_pagedown_files/paged-0.14/js/hooks.js

@@ -1,554 +1,0 @@
-// Hooks for paged.js
-{
-  // Utils
-  let pandocMeta, pandocMetaToString;
-  {
-    let el = document.getElementById('pandoc-meta');
-    pandocMeta = el ? JSON.parse(el.firstChild.data) : {};
-  }
-
-  pandocMetaToString = meta => {
-    let el = document.createElement('div');
-    el.innerHTML = meta;
-    return el.innerText;
-  };
-
-  let isString = value => {
-    return typeof value === 'string' || value instanceof String;
-  };
-
-  let isArray = value => {
-    return value && typeof value === 'object' && value.constructor === Array;
-  };
-
-  // This hook is an attempt to fix https://github.com/rstudio/pagedown/issues/131
-  // Sometimes, the {break-after: avoid;} declaration applied on headers
-  // lead to duplicated headers. I hate this bug.
-  // This is linked to the way the HTML source is written
-  // When we have the \n character like this: <div>\n<h1>...</h1>
-  // the header may be duplicated.
-  // But, if we have <div><h1>...</h1> without any \n, the problem disappear
-  // I think this handler can fix most of cases
-  // Obviously, we cannot suppress all the \n in the HTML document
-  // because carriage returns are important in <pre> elements.
-  // Tested with Chrome 76.0.3809.100/Windows
-  Paged.registerHandlers(class extends Paged.Handler {
-    constructor(chunker, polisher, caller) {
-      super(chunker, polisher, caller);
-      this.carriageReturn = String.fromCharCode(10);
-    }
-
-    checkNode(node) {
-      if (!node) return;
-      if (node.nodeType !== 3) return;
-      if (node.textContent === this.carriageReturn) {
-        node.remove();
-      }
-    }
-
-    afterParsed(parsed) {
-      let template = document.querySelector('template').content;
-      const breakAfterAvoidElements = template.querySelectorAll('[data-break-after="avoid"], [data-break-before="avoid"]');
-      for (let el of breakAfterAvoidElements) {
-        this.checkNode(el.previousSibling);
-        this.checkNode(el.nextSibling);
-      }
-    }
-  });
-
-  // This hook creates a list of abbreviations
-  // Note: we also could implement this feature using a Pandoc filter
-  Paged.registerHandlers(class extends Paged.Handler {
-    constructor(chunker, polisher, caller) {
-      super(chunker, polisher, caller);
-    }
-    beforeParsed(content) {
-      // Find the abbreviation nodes
[…]

### 2021-06-21 paper_pagedown.Rmd
modified rmd template.

@@
     self_contained: true
-    css: ['jss.css', 'jss-fonts.css', 'jss-page.css']
+    css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
 # uncomment this line to produce HTML and PDF in RStudio:

### 2021-06-21 paper_pagedown.Rmd
modified rmd template.

@@
 I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.).
-I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).'
+I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
 # at least one keyword must be supplied
@@
 paged-footnotes: true
-bibliography: index.bib
+bibliography: references.bib
 output: 
@@
 install.packages(c("rmarkdown", "knitr", "kableExtra",
-                   "stargazer", "plotly", "knitr"))
+                   "stargazer", "modelsummary", "plotly", "knitr"))
 ```
@@
 
-https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html
+The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R.
+Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
+This makes ist especially easzz to embed them in dznmaic documents.
+
+Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.
+
+In this template we demonstrate an example for modelsummary's `datasummary` function. `Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.
+
+### Summarize numeric variables with modelsummary
+
+Table \@ref(tab:tab4) shows a summary table for numeric variables.
+
+```{r tab4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+library(modelsummary)
+
+mtcars_df <- mtcars
+mtcars_df$vs <- as.logical(mtcars_df$vs)
+mtcars_df$cyl <- as.factor(mtcars_df$cyl)
+datasummary_skim(mtcars, type="numeric", output="html", histogram=T)
+```
+
+### Summarize categorical variables with modelsummary
+
+Table \@ref(tab:tab5) shows a summary table for categorical variables.
+
+```{r tab5, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+datasummary_skim(mtcars_df, type="categorical", output="html")
+```
 
@@
 
-html!
-
 # Compiling
@@
 - use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if you want your html based web page to be (additionally to the html!) printed to be PDF (the PDF will be stored in your working directory)
-- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.
+- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
 You can simply call the function `xaringan::inf_mr()` (within your console).
@@
 - If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio.
-Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.
+Use instead the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
 
@@
 
-```{r, include=FALSE}
-knitr::write_bib(c(.packages(), 'pagedown', 'xaringan'), 'index.bib')
-```
+::: {#refs}
 
-# Online appendix
+:::
+
+<!--  to generate a citation entry for BibTeX, you can pass the returned object of citation() to toBibtex(), e.g.
+toBibtex(citation("xaringan")).
+Then  copy the output to a .bib file and add a unique citation key -->
+<!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->
+
+# Online appendix {-}
 

### 2021-06-21 paper_pagedown.Rmd
changed font size and left aligned titles.
modified rmd template.

@@
 date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
-abstract: 'The present paper provides a template for a reproducible scientific paper written in R Markdown.
-Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document.
-The underlying files which produce this document can be downloaded [here](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
-I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.).
-I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
+abstract: 'The present paper provides a template for a reproducible scientific paper created with R pagedown.
+Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
+The underlying files which produce this document can be downloaded^[https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing].
+Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
+We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and R pagedown.
+We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
 # at least one keyword must be supplied
@@
     css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
-# uncomment this line to produce HTML and PDF in RStudio:
+    csl: american-sociological-association.csl
+# uncomment the next line if you only want to knit to HTML and not PDF
 knit: pagedown::chrome_print
 ---
-
-\clearpage
-
-\renewcommand{\baselinestretch}{0.5}\normalsize
-\tableofcontents
-\renewcommand{\baselinestretch}{1.1}\normalsize
-
-\clearpage
 
@@
 Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].
+
 * **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
@@
 It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.
+
 * **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
+
 * **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file.
@@
 
-\newpage
+Formatting texts as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
+The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.
+
+The fairly new `pagedown` R package takes a completely new approach.
+While the main purpose of pagedown is to create high-quality PDFs, the idea is to take advantage of modern web technologies (HTML/JSS/Javascript) with which one can design web pages and eventually print those to PDF.
+
+While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need.
+Additionally, pagedown documents are based on R Markdown.
+
 # Prerequesites
@@
 We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown.
-If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416).
-Also, there is an overview and template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).
+If you don't know what R Markdown is there are many great resources (e.g.
+watch this [short video](https://vimeo.com/178485416)).
+Also, there is a template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).
 The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.
@@
 Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents.
-For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing R's pagedown given by Yihui Xie (who in addition to Romain Lesur developed R's pagedown package).
+For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing `pagedown` given by Yihui Xie (who in addition to Romain Lesur developed the `pagedown` package).
 If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).
@@
 * ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style.
-If however you want to modify single aspects or create your own template, you will need at least some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
+If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
 
-\newpage
-# Basics: Input and output files
+# Basics: Input files, output files and the YAML header
 
@@
 
-Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line 18-22) to see how to specifiy these files.
+Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
 Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom css and html.
@@
 
-```{r, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 library(stargazer)

### 2021-06-22 paper_pagedown.Rmd
modified rmd template.

@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache = FALSE)
+# Use chache = TRUE if you want to speed up compilation
+# A function to allow for showing some of the inline code
+rinline <- function(code){
+  html <- '<code  class="r">``` `r CODE` ```</code>'
+  sub("CODE", code, html)
+}
+```
+
 \newpage
@@
 
-Formatting texts as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
+Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
 The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.
@@
 Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
-Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom css and html.
+Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom CSS and custom HTML.
 
 [Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder.
-Close R/Rstudio and directly open `paper.rmd` with RStudio.
+Close R/Rstudio and directly open `paper_pagedown.rmd` with RStudio.
 Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]
 
-Once you run/compile the `paper.rmd` file in Rstudio it creates mainly one output files:
+Once you run/compile the `paper.rmd` file in Rstudio it creates a output file called `paper_pagedown.html`.
 
-* `paper_pagedown.html` (the one you are reading right now)
+By using pagedown's `chrome_print` function in the YAML (line \#25) your html based web page will be printed to `paper_pagedown.pdf` (the one you are reading right now).
 
-By using pagedown's `chrome_print` function in the YAML (uncomment the line) your html based web page will be printed to be PDF (the PDF will be stored in your working directory)
-
-* `paper_pagedown.pdf`
+Both outputs will be saved in your working directory.
 
@@
 
-\newpage
 # Data
@@
 
-Generally, code is evaluated by inserting regular [R Markdown]{.proglang} blocks.
+Generally, code is evaluated by inserting regular `R Markdown` blocks.
 
@@
 
+## Putting your entire data into the .rmd file
+Applying the function `dput()` to an object gives you the code needed to reproduce that object.
+So you could paste that code into your `.rmd` file if you don't want to have extra data files.
+This makes sense were data files are small.
+
+```{r}
+dput(data)
+```
+
+You can then insert the dput output in your `.rmd` as below.
+
+```{r eval=FALSE, include=TRUE}
+data <- structure(list(X = 1:50, speed = c(4L, 4L, 7L, 7L, 8L, 9L, 10L, 
+10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, 
+14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, 
+18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, 
+24L, 24L, 24L, 25L), dist = c(2L, 10L, 4L, 22L, 16L, 10L, 18L, 
+26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, 
+36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, 
+76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, 
+92L, 93L, 120L, 85L)), 
+class = "data.frame", row.names = c(NA, 
+-50L))
+```
+
 # Tables {#sec:tables}
@@
 
+I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
+
 ### Summary table with stargazer
 
-Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
+Table \@ref(tab:summary) shows summary stats of your data.
 
-```{r, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
 library(stargazer)
 stargazer(cars, 
-          label="tab1", 
+          title = "Summary table with stargazer",
+          label="tab:summary", 
           table.placement = "H", 
@@
 
+Reproduction reaches new heights when you work with inline code.
+For instance, you can automatize the display of certain coefficients within the text.
+An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above.
+`r rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `r round(coef(model1)[2], 2)`.
+Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.
+Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects.
+For instance, imagine a reviewer asks you to omit certain observations from your sample.
+You can simply do so in the beginning of your code and push play subsequently..
+at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.
+Researchers often avoid referring to results in-text etc.
+because you easily forget to change them when revising a manuscript.
+At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it.
+Inline code allows you to do just that.
+R Markdown allows you to that do so in a reproducible and automatized manner.
+
 # Figures
@@
 
-To view your paper, pagedown requires a web server (since it is based von paged.js)^(open-source library to paginate content in the browser).
+To view your paper, pagedown requires a web server (since it is based von paged.js)^[open-source library to paginate content in the browser].
 By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.
 
-There are several options, depending on your intention.
+There are several options, depending on your intention:
 
-- click on the `Knit` button which by default will provide a html document in the RStudio viewer pane (the html will be stored in your working directory)
-- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if you want your html based web page to be (additionally to the html!) printed to be PDF (the PDF will be stored in your working directory)
+- click on the `Knit` button which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
+- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
 - to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.

### 2021-06-22 paper.Rmd
Added osf link.

@@
 ---
-title: |
-  |
-  | 
-  | \vspace{1cm}Writing a reproducible paper
-  | in Pagedown*^[*Corresponding address: mail@paulcbauer.eu. Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]\vspace{0.5cm}
-  |
-author: |
-  | Paul C. Bauer & Camille Landesvatter
-  | Mannheim Centre for European Social Research
-date: |
-  |       
-  | Based on: https://osf.io/q395s
-  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
-  |
-  | Feedback? Errors?
-  | mail@paulcbauer.eu
-  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
-linestretch: 1.2
-colorlinks: true
-abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown.
-Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document.
-The underlying files which produce this document can be downloaded [here](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
-I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.).
-I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
+title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]"
+shorttitle: "Writing a reproducible paper with R Markdown and Pagedown"
+author:
+  - name: Paul C. Bauer
+    affiliation: Mannheim Centre for European Social Research
+    email: <mail@paulcbauer.eu>
+  - name: Camille Landesvatter
+    affiliation: Mannheim Centre for European Social Research
+    email: <camille.landesvatter@mzes.uni-mannheim.de>
+date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
+download-link: 'Download: https://osf.io/k8jhx'
+abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
+The underlying files which produce this document can be downloaded^[https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing].
+Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
+We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and R pagedown.
+We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
+# at least one keyword must be supplied
+keywords: [open science, transparency, replication, R, markdown, pagedown]
+links-to-footnotes: true
+paged-footnotes: true
 bibliography: references.bib
+output: 
+  pagedown::jss_paged:
+    template: wp_paged.html
+    self_contained: true
+    css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
     csl: american-sociological-association.csl
-output:
-  bookdown::pdf_document2:
-    toc: no
-    keep_tex: true
-mainfont: Times New Roman
-sansfont: Times New Roman
-fontsize: 12pt
-link-citations: true
-documentclass: article
-geometry: margin=1in
-always_allow_html: yes
+# uncomment the next line if you only want to knit to HTML and not PDF
+knit: pagedown::chrome_print
 ---
@@
 # Use chache = TRUE if you want to speed up compilation
-
 # A function to allow for showing some of the inline code
@@
 
-\clearpage
+\newpage
+# Why reproducible research (in R)?
 
-\renewcommand{\baselinestretch}{0.5}\normalsize
-\tableofcontents
-\renewcommand{\baselinestretch}{1.1}\normalsize
-
-\clearpage
-
-# Why reproducible research (in R)?
 Some arguments...
@@
 
+# Why Pagedown?
+
+Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
+The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.
+
+The fairly new `pagedown` R package takes a completely new approach.
+While the main purpose of pagedown is to create high-quality PDFs, the idea is to take advantage of modern web technologies (HTML/JSS/Javascript) with which one can design web pages and eventually print those to PDF.
+
+While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need.
+Additionally, pagedown documents are based on R Markdown.
+
 # Prerequesites
-I assume that you are using R on a day-to-day basis.
-You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown.
-If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416).
+
+We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown.
+If you don't know what R Markdown is there are many great resources (e.g.
+watch this [short video](https://vimeo.com/178485416)).
+Also, there is a template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).
+The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.
+
+Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents.
+For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing `pagedown` given by Yihui Xie (who in addition to Romain Lesur developed the `pagedown` package).
+If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).
+
 Then...
@@
 * ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
-* ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex].
+* ...install the "pagedown"-package from github using the code below.
 ```{r eval=FALSE, include=TRUE}
-install.packages(c('tinytex', 'rmarkdown'))
-tinytex::install_tinytex()
+remotes::install_github('rstudio/pagedown')
 ```
-* ...install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
+* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
 ```{r eval=FALSE, include=TRUE}
 install.packages(c("rmarkdown", "knitr", "kableExtra",
-                   "stargazer", "plotly", "knitr",
-                   "bookdown"))
+                   "stargazer", "modelsummary", "plotly", "knitr"))
 ```
-* ...download the 5 input files I created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
+* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
 Ignore the other files.
+* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
+* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
+* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
+* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style.
+If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
 
-* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
+# Basics: Input files, output files and the YAML header
 
-# Basics: Input and output files
-All the files you need to produce the present PDF file are the input files...
+All the files you need to produce the present PDF file are:
 
-* ...a `paper.rmd` file (the underlying R Markdown file).
-* ...a `references.bib` file (the bibliography).
+1. the input files:
+
+* `paper.rmd` (the underlying R Markdown file).
+* `references.bib` (the bibliography).
     + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.rmd` file.
-* ...a `data.csv` file (some raw data).
-* ...
-a `american-sociological-association.csl` file that defines the style of your bibliography.^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]
+* `data.csv` (some raw data).
+* `american-sociological-association.csl` (defines the style of your bibliography).^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]
+
+2. the "styling" files:
+
+Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document.
+With using our templates, you will create a document that has the "look" of a working paper.
+
+* `wp_paged.html`
+* `wp.css`
+* `wp-fonts.css`
+* `wp-pages.css`
+
+Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
+Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom CSS and custom HTML.
 
 [Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder.
-Close R/Rstudio and directly open `paper.rmd` with RStudio.
+Close R/Rstudio and directly open `paper_pagedown.rmd` with RStudio.
 Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]
 
-Once you run/compile the `paper.rmd` file in Rstudio it creates mainly two output files:
+Once you run/compile the `paper.rmd` file in Rstudio it creates a output file called `paper_pagedown.html`.
 
-* `paper.tex`
-* `paper.pdf` (the one you are reading right now)
+By using pagedown's `chrome_print` function in the YAML (line \#25) your html based web page will be printed to `paper_pagedown.pdf` (the one you are reading right now).
 
-In addition, there may be files that you generate and store locally in the folder during the compilation process. This is the case for some of the Plotly graphs below.
+Both outputs will be saved in your working directory.
 
-Ideally, we can simply provide others with a `zip` folder that contains both our input files and our output files.
-Then it's possible to reproduce the process from managing/analyzing some raw data to producing the final scientific article.
-
-Below we always display the R code in the chunks that produce the output.
-In your paper you will normally only present outputs (e.g., tables, figures etc.) by choosing the chunk option "Show output only" in R Studio.
-The chunk commands itself are not displayed but they do matter for referencing etc.
-So simply orient yourself at the underlying `paper.rmd` file.
-
+\newpage
 # Referencing within your document
@@
 # Software versioning
+
 Software changes and gets updated, especially with an active developer community like that of R.
@@
 ## Import
-```{r, echo=TRUE}
+
+Generally, code is evaluated by inserting regular `R Markdown` blocks.
+
+```{r}
+x <- 1:10
+x
+```
+
+Below we import an exemplary dataset ([download](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing)).
+
+```{r, echo=T, results="raw"}
 data <- read.csv("data.csv")
@@
 # Tables {#sec:tables}
+
 Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better.
@@
 
-Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
+I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
+
+### Summary table with stargazer
+
+Table \@ref(tab:summary) shows summary stats of your data.
 
@@
           title = "Summary table with stargazer",
-          label="tab1", 
+          label="tab:summary", 
           table.placement = "H", 
+          type="html",
           header=FALSE)
@@
 
+### Regression table with stargazer
+
 Table \@ref(tab2) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
 
-```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 library(stargazer)
@@
 model3 <- lm(dist ~ speed, data = cars)
+
 stargazer(model1, model2, model3,
-          title = "Regression table with stargazer",
           label="tab2",
@@
           column.labels = c("M1", "M2", "M3"),
+          type="html",
           model.numbers = FALSE,
@@
 Another great function is `kable()` (`knitr` package) in combination with `kableExtra`.
-Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.] Again you can modify so many things in both the `kable()` and the `kable_styling()` function.
-See [this overview](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) of all the kable stylings that are possible provided by the package author himself.
+Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.]
 
-```{r tab3, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+```{r tab3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 library(knitr)
 library(kableExtra)
-kable(cars[1:10,], row.names = TRUE, 
+
+kable(mtcars[1:10,], row.names = TRUE, 
       caption = 'Table with kable() and kablestyling()', 
-      format = "latex", booktabs = T) %>%
+      format = "html", booktabs = T) %>%
         kable_styling(full_width = T, 
@@
                       font_size = 10)
+
+```
+
+## modelsummary
+
+The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R.
+Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
+This makes ist especially easzz to embed them in dznmaic documents.
+
+Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.
+
+In this template we demonstrate an example for modelsummary's `datasummary` function. `Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.
+
+### Summarize numeric variables with modelsummary
+
+Table \@ref(tab:tab4) shows a summary table for numeric variables.
+
+```{r tab4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+library(modelsummary)
+
+mtcars_df <- mtcars
+mtcars_df$vs <- as.logical(mtcars_df$vs)
+mtcars_df$cyl <- as.factor(mtcars_df$cyl)
+datasummary_skim(mtcars, type="numeric", output="html", histogram=T)
+```
+
+### Summarize categorical variables with modelsummary
+
+Table \@ref(tab:tab5) shows a summary table for categorical variables.
+
+```{r tab5, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+datasummary_skim(mtcars_df, type="categorical", output="html")
 ```
@@
 # Inline code & results
+
 Reproduction reaches new heights when you work with inline code.
@@
 
+## Interactive graphs
+
+# Compiling
+
+To view your paper, pagedown requires a web server (since it is based von paged.js)^[open-source library to paginate content in the browser].
+By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.
+
+There are several options, depending on your intention:
+
+- click on the `Knit` button which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
+- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
+- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
+You can simply call the function `xaringan::inf_mr()` (within your console).
+This will launch a local web server via the **servr** package (Xie 2021a) and display your pages in the RStudio viewer.
+Each time you save your document (_Ctrl+S_) xaringan updates your pages in the viewer.
+
+- If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio.
+Use instead the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
+
 # Good practices
@@
     + "`recoding-...`" for chunks in which data is recoded
-* Ideally give chunks the labels of the figures/tables in the final publication, e.g., "`fig-1`" is the name of the chunk that produces Figure 1 in the final paper.
 * Use "really" informative variable names:
@@
 
-# References
+\newpage
 
-\linespread{1}
+# References {-}
 
@@
 
-\clearpage
+<!--  to generate a citation entry for BibTeX, you can pass the returned object of citation() to toBibtex(), e.g.
+toBibtex(citation("xaringan")).
+Then  copy the output to a .bib file and add a unique citation key -->
+<!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->
 
-\appendix
-\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
-
-\renewcommand{\thesection}{A}
-
-\setcounter{page}{1}
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\renewcommand{\figurename}{Table}
-
-\setcounter{figure}{0}
-\renewcommand\thefigure{A\arabic{figure}}
-\renewcommand{\figurename}{Figure}
-
-\clearpage
-\pagenumbering{gobble}
-
-\vspace*{7cm}
-
-\begin{center}
-\begin{huge}
-Online Appendix
-\end{huge}
-\end{center}
-\vspace{3cm}
-
-\clearpage
-\pagenumbering{arabic}
-
-# Online appendix
+# Online appendix {-}
 

### 2021-06-22 paper.Rmd
Various small edits.

@@
 download-link: 'Download: https://osf.io/k8jhx'
-abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
+abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
 The underlying files which produce this document can be downloaded^[https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing].
@@
 ```
-* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
+* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
 ```{r eval=FALSE, include=TRUE}
 install.packages(c("rmarkdown", "knitr", "kableExtra",
-                   "stargazer", "modelsummary", "plotly", "knitr"))
+                   "stargazer", "modelsummary", "knitr"))
 ```
@@
 
-## modelsummary
+## Summary tables with modelsummary
 
@@
 
-### Summarize numeric variables with modelsummary
+### Summarize numeric variables
 
@@
 library(modelsummary)
-
-mtcars_df <- mtcars
-mtcars_df$vs <- as.logical(mtcars_df$vs)
-mtcars_df$cyl <- as.factor(mtcars_df$cyl)
 datasummary_skim(mtcars, type="numeric", output="html", histogram=T)
@@
 
-### Summarize categorical variables with modelsummary
+### Summarize categorical variables
 
@@
 ```{r tab5, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
-datasummary_skim(mtcars_df, type="categorical", output="html")
+# Create categorical variables
+mtcars$vs_cat <- as.logical(mtcars$vs)
+mtcars$cyl_cat <- as.factor(mtcars$cyl)
+datasummary_skim(mtcars, type="categorical", output="html")
 ```
@@
 
-# Compiling
+# Compiling the document
 
@@
 
-- click on the `Knit` button which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
+- click on the `Knit` button in R Studio which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
 - use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
@@
 
-# Good practices
-Every researcher has his own optimized setup. Currently I would recommend the following:
+# Good practices for reproducibility
+Every researcher has his own optimized setup. Currently we would recommend the following:
 
 * Keep all files of your project (that matter for producing the PDF) in one folder without subfolders.
-You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).
+You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).^[Another good folder setup would be to store all files needed as input files for the R Markdown manuscript in a subfolder called "input" and all output files that are produced apart from paper.html and paper.pdf in a subfolder called "output".]
 * Make sure that filenames have a logic to them.
@@
     + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
-        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher. 
+        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher in 50 years. 
     + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
-* Use unique identifiers in the final document:
-    + e.g., name the models you estimate "M1", "M2" etc.
-    + These unique names should also appear in the published paper.
-    + Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code...
+* Use unique identifiers in the final R Markdown document paper.rmd that you upload:
+    + Think of someone who wants to produce Figure 1/Model 1/Table 1 in your paper but doesn't find it in your code...
+        + Name the chunks "fig-1", "fig-2" as the are named in the published paper.
+        + Name the chunks that produce tables "table-1", "table-2" etc. as they are named in the published paper.
+        + Name your statistical models in your R code "M1", "M2" as they are named in the published paper.
 
@@
 
+\newpage
+
 # Online appendix {-}

### 2021-06-22 paper.Rmd
Smaller edits.

@@
 The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.
+However, if you have spent hours and hours debugging latex code (or getting it to run) you may be on the lookout for something new.
 
@@
 Additionally, pagedown documents are based on R Markdown.
+In our view, Pagedown and the underlying technology may replace Latex in the long run.
 
@@
 We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown.
-If you don't know what R Markdown is there are many great resources (e.g.
+If you don't know what R Markdown is there are many great resources that you should use (e.g.
 watch this [short video](https://vimeo.com/178485416)).
-Also, there is a template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).
-The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.
+An older template [see @Bauer2018-dl; https://osf.io/q395s] on which this newer template is based, may provide a quick entry point to writing a reproducible with R Markdown and Latex.
 
@@
 * ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
-* ...install the "pagedown"-package from github using the code below.
+* ...install the "pagedown"-package from github using the code below [@Xie2021-bi; @Xie2021-ls].
 ```{r eval=FALSE, include=TRUE}

### 2021-06-22 paper.Rmd
added background color to R Code Chunks.

@@
 # uncomment the next line if you only want to knit to HTML and not PDF
-knit: pagedown::chrome_print
+#knit: pagedown::chrome_print
 ---

### 2021-06-22 paper.Rmd
changed fonts.
Add cross-referenced regression table.

--- 2021-06-22 paper.Rmd
@@
 abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be downloaded^[https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing].
+The underlying files which produce this document can be [downloaded](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
 Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
@@
 # Why reproducible research (in R)?
-
+<a href="https://www.w3schools.com/">Visit W3Schools.com!</a>
 Some arguments...
@@
 * ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
-* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
+* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Pagedown](https://pagedown.rbind.io/).
 * ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style.
@@
 
-\newpage
 # Referencing within your document
@@
 
-## stargazer(): Summary and regression tables
-
-I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
-
-### Summary table with stargazer
-
-Table \@ref(tab:summary) shows summary stats of your data.
-
-```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
-library(stargazer)
-stargazer(cars, 
-          title = "Summary table with stargazer",
-          label="tab:summary", 
-          table.placement = "H", 
-          type="html",
-          header=FALSE)
-```
-
-### Regression table with stargazer
-
-Table \@ref(tab2) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
-
-```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
-library(stargazer)
-model1 <- lm(speed ~ dist, data = cars)
-model2 <- lm(speed ~ dist, data = cars)
-model3 <- lm(dist ~ speed, data = cars)
-
-stargazer(model1, model2, model3,
-          label="tab2",
-          table.placement = "H", 
-          column.labels = c("M1", "M2", "M3"),
-          type="html",
-          model.numbers = FALSE,
-          header=FALSE)
-```
-
 ## kable() and kable_styling()
 
-Another great function is `kable()` (`knitr` package) in combination with `kableExtra`.
-Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.]
+A great function is `kable()` (`knitr` package) in combination with `kableExtra`.
+Table \@ref(tab:table-2) provides an example.
+To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:table-2´ and would reference it by adding "`Table \@ref(tab:table-2)`" in your text.
 
-```{r tab3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r table-2, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 library(knitr)
@@
 
-## Summary tables with modelsummary
+## Tables with modelsummary
 
@@
 This makes ist especially easzz to embed them in dznmaic documents.
-
 Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.
-
-In this template we demonstrate an example for modelsummary's `datasummary` function. `Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.
+In this template we demonstrate an example for modelsummary's `datasummary` function.
+`Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.
 
@@
 
-Table \@ref(tab:tab4) shows a summary table for numeric variables.
+Table \@ref(tab:table-3) shows a summary table for numeric variables.
 
-```{r tab4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r table-3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 library(modelsummary)
-datasummary_skim(mtcars, type="numeric", output="html", histogram=T)
+library(kableExtra)
+options(modelsummary_html = 'kableExtra')
+datasummary_skim(mtcars, type="numeric", output="html", histogram=T) %>%
+    kable_styling()
 ```
@@
 
-Table \@ref(tab:tab5) shows a summary table for categorical variables.
+Table \@ref(tab:table-4) shows a summary table for categorical variables.
 
-```{r tab5, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r table-4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
 # Create categorical variables
@@
 datasummary_skim(mtcars, type="categorical", output="html")
+```
+
+## Regression table
+
+Table \@ref(tab:table-5) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
+
+```{r table-5, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+library(modelsummary)
+M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
+M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
+M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
+models <- list("M1" = M1, "M2" =  M2, "M3" = M3)
+
+modelsummary(models, 
+         title = 'A regression table.') %>%
+    kable_styling()
 ```
@@
 An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above.
-`r rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `r round(coef(model1)[2], 2)`.
+`r rinline("round(coef(M1)[2], 2)")` will insert the coefficient as follows: `r round(coef(M1)[2], 2)`.
 Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.

### 2021-06-22 paper.Rmd
Repaired table cross-references

--- 2021-06-22 paper.Rmd
@@
 
-```{r table-3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r table-3, echo=TRUE}
 library(modelsummary)
-library(kableExtra)
-options(modelsummary_html = 'kableExtra')
-datasummary_skim(mtcars, type="numeric", output="html", histogram=T) %>%
-    kable_styling()
+datasummary_skim(mtcars, 
+                 type="numeric", 
+                 histogram=T, 
+                 title = "Summary: Numeric variables")
 ```
@@
 
-```{r table-4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
+```{r table-4, echo=TRUE}
 # Create categorical variables
@@
 mtcars$cyl_cat <- as.factor(mtcars$cyl)
-datasummary_skim(mtcars, type="categorical", output="html")
+datasummary_skim(mtcars, 
+                 type="categorical", 
+                 title = "Summary: Categorical variables")
 ```
@@
 
-```{r table-5, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+```{r table-5, echo=TRUE}
 library(modelsummary)
@@
 
-modelsummary(models, 
-         title = 'A regression table.') %>%
-    kable_styling()
+modelsummary(models, title = 'A regression table.')
 ```

### 2021-06-23 paper.Rmd
Small edits.
Deleted faulty text.
Updated inline links.

--- 2021-06-22 paper.Rmd
@@
 abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
+The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
 Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
@@
 keywords: [open science, transparency, replication, R, markdown, pagedown]
-links-to-footnotes: true
+links-to-footnotes: false
 paged-footnotes: true
@@
 # Why reproducible research (in R)?
-<a href="https://www.w3schools.com/">Visit W3Schools.com!</a>
 Some arguments...
@@
 ```
-* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing).
+* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
 Ignore the other files.
@@
 
-[Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder.
+[Download these files](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing) and save them into a folder.
 Close R/Rstudio and directly open `paper_pagedown.rmd` with RStudio.
@@
 
-## kable() and kable_styling()
+## Tables with kable() and kable_styling()
 
@@
 Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
-This makes ist especially easzz to embed them in dznmaic documents.
+This makes ist especially easy to embed them in dynamic documents.
 Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.
@@
 ```
-
-## Interactive graphs
 

### 2021-06-23 paper.Rmd
first commit plotly graph.

@@
 output: 
+  html_document: #for plotly
   pagedown::jss_paged:
@@
 ```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
+library(ggplot2)
+
 mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
-library(ggplot2)
+
 ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
@@
 
+## Interactive graphs
+
+```{r , echo=FALSE, out.width='100%', message=FALSE, warning=FALSE}
+
+library(plotly)
+
+mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
+mtcars %>%
+  plot_ly(x=~wt,y=~mpg,color=~cyl) %>%
+  add_markers() %>% 
+  layout(xaxis=list(title="Weight (lb/1000)"), yaxis=list(title="Miles/(US) gallon"))
+```
+
 # Compiling the document
 
-To view your paper, pagedown requires a web server (since it is based von paged.js)^[open-source library to paginate content in the browser].
+To view your paper, pagedown requires a web server (since it is based on paged.js)^[open-source library to paginate content in the browser].
 By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.

### 2021-06-23 paper.Rmd
first commit plotly graph.

@@
 # uncomment the next line if you only want to knit to HTML and not PDF
-#knit: pagedown::chrome_print
+knit: pagedown::chrome_print
 ---

### 2021-06-23 paper.Rmd
removed plotly graph.

@@
 output: 
-  html_document: #for plotly
   pagedown::jss_paged:
@@
 
-## Interactive graphs
-
-```{r , echo=FALSE, out.width='100%', message=FALSE, warning=FALSE}
-
-library(plotly)
-
-mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
-mtcars %>%
-  plot_ly(x=~wt,y=~mpg,color=~cyl) %>%
-  add_markers() %>% 
-  layout(xaxis=list(title="Weight (lb/1000)"), yaxis=list(title="Miles/(US) gallon"))
-```
-
 # Compiling the document

### 2021-06-23 paper.Rmd
Small edits to text

@@
 Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document.
-With using our templates, you will create a document that has the "look" of a working paper.
+With using our templates, you will create a document that has the "look" of a working paper (we based our files on the [jss_paged pagedown format](https://github.com/rstudio/pagedown#journal-of-statistical-software-article-pagedownjss_paged)).
 
-* `wp_paged.html`
+* `wp_paged.html` (based on `jss_paged.html`)
 * `wp.css`
@@
 
-# Figures
+# Graphs
 

### 2021-06-23 paper.Rmd
Small text edits.
modified tables.

--- 2021-06-23 paper.Rmd
@@
 I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \@ref(sec:rsessioninfo).
-This will make sure you always provide the package versions that you used in the last compilation of your paper.
+This will make sure you always inform the reader about the package versions your relied on in your paper.
 For more advanced tools see [packrat](https://rstudio.github.io/packrat/).
@@
 mtcars$cyl_cat <- as.factor(mtcars$cyl)
-datasummary_skim(mtcars, 
+
+library(flextable)
+tab_cat <- datasummary_skim(mtcars, 
                             type="categorical", 
-                 title = "Summary: Categorical variables")
+                            title = "Summary: Categorical variables",
+                            output = 'flextable')
+
+# additionally we want to change the font, fontsize and spacing
+library("gdtools")
+
+library(dplyr)
+tab_cat <- tab_cat %>%
+  font(fontname="Times New Roman", part="header") %>%
+  font(fontname="Times New Roman", j=1:4) %>% 
+  fontsize(size=12, part="header") %>%
+  fontsize(size=10, j=1:4) %>% 
+  line_spacing(space = 0.3, part = "all")
+
+tab_cat
 ```
@@
 
-modelsummary(models, title = 'A regression table.')
+tab_reg <-modelsummary(models, title = 'A regression table.', output = 'flextable')
+
+# additionally we want to change the font, font size and spacing
+library("gdtools")
+
+tab_reg <- tab_reg %>%
+  font(fontname="Times New Roman", part="header") %>%
+  font(fontname="Times New Roman", j=1:4) %>% 
+  fontsize(size=10, part="header") %>%
+  fontsize(size=10, j=1:4) %>% 
+  line_spacing(space = 0.3, part = "all")
+
+tab_reg
+
 ```

### 2021-06-23 paper.Rmd
small edits.

@@
 
-```{r table-3, echo=TRUE}
+```{r table-3, echo=TRUE, message=FALSE, warning=FALSE}
 library(modelsummary)
@@
 
-```{r table-4, echo=TRUE}
+```{r table-4, echo=TRUE, message=FALSE, warning=FALSE}
 # Create categorical variables
@@
 
-```{r table-5, echo=TRUE}
+```{r table-5, echo=TRUE, message=FALSE, warning=FALSE}
 library(modelsummary)

### 2021-06-23 paper.Rmd
Changed data from mtcars to swiss.

@@
 Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
-We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and R pagedown.
-We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
+We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and Pagedown.
+We intend to update this file when we discover more convenient code.
+You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/.'
 # at least one keyword must be supplied
@@
 In our view, Pagedown and the underlying technology may replace Latex in the long run.
+In the near future it should also be possible to produce a PDF with static graph and an equivalent html with interactive graps (see dicsussion [here](https://github.com/rstudio/pagedown/pull/87)).
 
@@
 
-Below we import an exemplary dataset ([download](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing)).
+Below we import an exemplary dataset (you can find `data.csv` in the folder with the other files).
 
@@
 ```{r}
-dput(data)
+dput(data[1:5,]) # here we only take a subset
 ```
@@
 ```{r eval=FALSE, include=TRUE}
-data <- structure(list(X = 1:50, speed = c(4L, 4L, 7L, 7L, 8L, 9L, 10L, 
-10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, 
-14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, 
-18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, 
-24L, 24L, 24L, 25L), dist = c(2L, 10L, 4L, 22L, 16L, 10L, 18L, 
-26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, 
-36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, 
-76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, 
-92L, 93L, 120L, 85L)), 
-class = "data.frame", row.names = c(NA, 
--50L))
+data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
+45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
+), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
+93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
+20.6)), class = "data.frame", row.names = c(NA, -5L))
 ```
@@
 
-Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better.
-Examples that you may use are shown below.
-The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.
+Producing good tables and referencing these tables within a R Markdown document has been a hassle but got much better. Examples that you may use are shown below.
 
@@
 
-kable(mtcars[1:10,], row.names = TRUE, 
+kable(swiss[1:10,], row.names = TRUE, 
       caption = 'Table with kable() and kablestyling()', 
@@
 library(modelsummary)
-datasummary_skim(mtcars, 
+datasummary_skim(swiss, 
                  type="numeric", 
@@
 # Create categorical variables
-mtcars$vs_cat <- as.logical(mtcars$vs)
-mtcars$cyl_cat <- as.factor(mtcars$cyl)
+swiss$Education_cat <- cut(swiss$Education, 
+                   breaks=c(-Inf, 6, 12, Inf), 
+                   labels=c("low","middle","high"))
+swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
+                   breaks=c(-Inf, 18.15, 21.70, Inf), 
+                   labels=c("low","middle","high"))
 
 library(flextable)
-tab_cat <- datasummary_skim(mtcars, 
+tab_cat <- datasummary_skim(swiss, 
                             type="categorical", 
@@
 ```{r fig-1, fig.align="center", fig.cap="Scatterplot of Speed and Distance", fig.pos="H", message=FALSE, warning=FALSE, paged.print=FALSE}
-plot(cars$speed, cars$dist)
+plot(swiss$Catholic, swiss$Fertility)
 ```
@@
 library(ggplot2)
-
-mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
-
-ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
-  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
-       shape="Number of \n Cylinders") + theme_classic()
+ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) + geom_point() +
+  labs(x="Agriculture", y = "Fertility", 
+       shape="Education") + theme_classic()
 ```

### 2021-06-23 paper.Rmd
Updated keywords

@@
 We intend to update this file when we discover more convenient code.
-You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/.'
+You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/).'
 # at least one keyword must be supplied
-keywords: [open science, transparency, replication, R, markdown, pagedown]
+keywords: [open science, transparency, replication, reproducible research, reproducibility, R, markdown, pagedown]
 links-to-footnotes: false

### 2022-01-28 paper.Rmd
Use gt for regression table.

@@
 install.packages(c("rmarkdown", "knitr", "kableExtra",
-                   "stargazer", "modelsummary", "knitr"))
+                   "stargazer", "modelsummary", "knitr", "gt"))
 ```
@@
 
-Table \@ref(tab:table-5) shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
+Table \@ref(tab:table-5) shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
 
@@
 
-tab_reg <-modelsummary(models, title = 'A regression table.', output = 'flextable')
-
+library(gt)
 # additionally we want to change the font, font size and spacing
-library("gdtools")
-
-tab_reg <- tab_reg %>%
-  font(fontname="Times New Roman", part="header") %>%
-  font(fontname="Times New Roman", j=1:4) %>% 
-  fontsize(size=10, part="header") %>%
-  fontsize(size=10, j=1:4) %>% 
-  line_spacing(space = 0.3, part = "all")
-
-tab_reg
+modelsummary(models,
+             title = 'Linear regression',
+             output = 'gt',
+             notes = "Notes: some notes...") %>%
+    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4) %>%
+  tab_options(
+    table.font.size = 10,
+    data_row.padding = px(1),
+    table.border.top.color = "white",
+    heading.border.bottom.color = "black",
+    row_group.border.top.color = "black",
+    row_group.border.bottom.color = "white",
+    table.border.bottom.color = "white",
+    column_labels.border.top.color = "black",
+    column_labels.border.bottom.color = "black",
+    table_body.border.bottom.color = "black",
+    table_body.hlines.color = "white"
+  )
 

### 2023-04-05 paper.Rmd
Update Links/URLs.

@@
 ---
-title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]"
+title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.de, camille.landesvatter@uni-mannheim.de]"
 shorttitle: "Writing a reproducible paper with R Markdown and Pagedown"
@@
     affiliation: Mannheim Centre for European Social Research
-    email: <mail@paulcbauer.eu>
+    email: <mail@paulcbauer.de>
   - name: Camille Landesvatter
     affiliation: Mannheim Centre for European Social Research
-    email: <camille.landesvatter@mzes.uni-mannheim.de>
+    email: <camille.landesvatter@uni-mannheim.de>
 date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
@@
 abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
+The underlying files which produce this document can be [downloaded here](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown) (click on Code -> Download ZIP).
 Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
@@
 ```
-* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
+* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [Github](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown) (click on Code -> Download ZIP).
 Ignore the other files.
@@
 
-[Download these files](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing) and save them into a folder.
+[Download these files](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown) (click on Code -> Download ZIP) and save them into a folder.
 Close R/Rstudio and directly open `paper_pagedown.rmd` with RStudio.
